---
title: "Sentiment Analysis"
author: "Alexandre Galiani Garmbis"
date: "17/09/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidytext)
```

## Lexicons

- *afinn* from Finn Årup Nielsen;
- *bing* from Bing Liu and collaborators, and;
- *nrc* from Saif Mohammad and Peter Turney.

```{r}
get_sentiments("afinn")
```


```{r}
(bing <- get_sentiments("bing"))
```

```{r}
nrc <- get_sentiments("nrc") 
nrc %>% count(sentiment)
```

## Twitter data

Tweets geocoded by US states:

```{r}
load("data/geocoded_tweets.rda")
tweets_bing <- geocoded_tweets %>% inner_join(bing)
tweets_bing %>% 
  group_by(state, sentiment) %>%
  summarize(freq = mean(freq)) %>%
  spread(sentiment, freq) %>%
  mutate(pos_ratio = positive/(positive + negative)) %>%
  arrange(desc(pos_ratio))
```

## Shakespeare data

### Transform the non-tidy text data to tidy text data
```{r}
load("data/shakespeare.rda")
tidy_shakespeare <- shakespeare %>%
  group_by(title) %>%
  mutate(linenumber = row_number()) %>%
  unnest_tokens(word, text) %>%
  ungroup()

tidy_shakespeare %>% 
  count(word, sort = TRUE)
```

### Implement sentiment analysis with the "bing" lexicon
```{r}
shakespeare_sentiment <- tidy_shakespeare %>%
  inner_join(get_sentiments("bing"))

word_counts <- shakespeare_sentiment %>%
  count(title, sentiment) 

word_counts %>%
  ggplot(aes(title, n, fill = sentiment)) +
  geom_col(position = "fill") +
  coord_flip()
```

### Top words by sentiment
```{r}
top_words <- word_counts %>%
  group_by(sentiment) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(word = reorder(word, n))

ggplot(top_words, aes(word, n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free") +  
  coord_flip()
```
Obs.: Wilt is not a negative word in this context, neither hero a positive (proper name).  

### Calculate a contribution for each word in each title
  
```{r}
sentiment_contributions <- tidy_shakespeare %>%
  # Count by title and word
  count(title, word, sort = TRUE) %>%
  # Implement sentiment analysis using the "afinn" lexicon
  inner_join(get_sentiments("afinn")) %>%
  # Group by title
  group_by(title) %>%
  # Calculate a contribution for each word in each title
  mutate(contribution = n * score / sum(n)) %>%
  anti_join(data_frame(word = c("no","yes","hero","wilt"))) %>%
  arrange(desc(contribution)) %>%
  filter(title == "The Tragedy of Macbeth")
    
sentiment_contributions
```

### Sentiment changes through the narrative arcs 

```{r}
tidy_shakespeare %>%
  inner_join(get_sentiments("bing")) %>%
  count(title, type, index = linenumber %/% 100, sentiment) %>%
  # Spread sentiment and n across multiple columns
  spread(sentiment, n, fill = 0) %>%
  # Use mutate to find net sentiment
  mutate(sentiment = positive - negative) %>%
  ggplot(aes(index, sentiment)) + geom_col(aes(fill = sentiment > 0)) +
  facet_wrap(~ type, scales = "free_x") +
  theme(legend.position = "none")
```

## Climate test data

```{r}
load("data/climate_text.rda")
climate_text %>% 
  unnest_tokens(word, text) %>%
  inner_join(get_sentiments("bing")) %>%
  ggplot(aes(station, fill = sentiment)) + 
  geom_bar(position = "fill") +
  coord_flip()
```
### Top 10 word for each sentiment (proper names excluded)
```{r}
climate_text %>% 
  unnest_tokens(word, text) %>%
  inner_join(get_sentiments("nrc"))  %>%
  anti_join(data_frame(word = c("change", "gore", "trump"))) %>%
  count(sentiment, word) %>%
  group_by(sentiment) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ sentiment, scales = "free") +
  coord_flip()
```

### Positiveness by time
```{r}
library(lubridate)

climate_text %>% 
  unnest_tokens(word, text) %>%
  mutate(date = floor_date(show_date, unit = "6 months")) %>%
  inner_join(get_sentiments("bing")) %>%
  count(station, date, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(positiveness = positive / (positive + negative)) %>%
  ggplot(aes(date, positiveness)) +
  geom_line() +
  geom_smooth(method = "lm", se = FALSE, lty = 2) +
  expand_limits(y = 0) +
  facet_wrap(~station, ncol = 1)
```

## Pop song lyrics

- rank, the rank a song achieved on the Billboard Year-End Hot 100,
- song, the song's title,
- artist, the artist who recorded the song,
- year, the year the song reached the given rank on the Billboard chart, and
- lyrics, the lyrics of the song.

```{r}
load("data/song_lyrics.rda")
tidy_lyrics <- song_lyrics %>% 
  unnest_tokens(word, lyrics)

totals <- tidy_lyrics %>%
  count(song) %>%
  rename(total_words = n)

lyric_sentiment <- tidy_lyrics %>%
  left_join(totals, by = "song") %>%
  inner_join(get_sentiments("nrc"))

lyric_sentiment %>%
  count(song, sentiment, sort = TRUE)
```

### What songs have the highest proportion of negative words?

```{r}
lyric_sentiment %>%
    count(song, sentiment, total_words) %>%
    mutate(percent = n / total_words) %>%
    filter(sentiment == "negative") %>%
    arrange(desc(percent))
```

### What songs have the highest proportion of positive words?

```{r}
lyric_sentiment %>%
    count(song, sentiment, total_words) %>%
    mutate(percent = n / total_words) %>%
    filter(sentiment == "positive") %>%
    arrange(desc(percent))
```

### How is positive sentiment changing over time?

```{r}
lyric_sentiment %>%
    filter(sentiment == "positive") %>%
    count(song, year, total_words) %>%
    ungroup() %>%
    mutate(percent = n / total_words,
           decade = 10 * floor(year / 10)) %>%
    ggplot(aes(as.factor(decade), percent, color = year)) +
    scale_y_log10() +
    geom_jitter(alpha = 0.1, width = 0.1)
```

```{r}
positive_by_year <- lyric_sentiment %>%
    filter(sentiment == "positive") %>%
    # Count by song, year, and total number of words
    count(song, year, total_words) %>%
    ungroup() %>%
    # Define a new column: percent
    mutate(percent = n / total_words)

# Fit a linear model with percent as the response and year as the predictor
model_positive <- lm(percent ~ year, data = positive_by_year)

# Use summary to see the results of the model fitting
summary(model_positive)
```

